---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rsgeo

This package is in very very very very very infantile state. This is an R package to add bindings to the rust geo crate. 

Load the package
```{r}
devtools::load_all()
```


Here is a performance challenge i am running into. 

Create geometries.

```{r}
# get geometry from sf
polys <- sfdep::guerry[["geometry"]] |>
  sf::st_cast("POLYGON")

# cast to rust
rs_polys <- rs_polygons(polys)
```


Benchmark one-to-one intersections.

```{r}
bench::mark(
  rust = intersect_poly_poly(rs_polys[[1]], rs_polys[[85]]),
  sf = sf::st_intersects(polys[[1]], polys[[85]]),
  check = FALSE
)
```

Rust is insanely fast. 

Do one-to-many it becomes super slow but still very memory efficient. 

```{r}
bench::mark(
  rust = intersect_poly_polys(rs_polys[[1]], rs_polys),
  sf = sf::st_intersects(polys[[1]], polys),
  check = FALSE
)
```

Rust implementation for `intersect_poly_poly()`. See line 353 in lib.rs.

```rust
fn intersect_poly_poly(lhs: Robj, rhs: Robj) -> Rbool {
    let xpoly: ExternalPtr<Polygon> = lhs.try_into().unwrap(); 
    let ypoly: ExternalPtr<Polygon> = rhs.try_into().unwrap(); 

    Rbool::from(xpoly.intersects(&*ypoly))

}
```

This is nice and fast. 

The implementation for `intersect_poly_polys()`. See line 383

```rust
fn intersect_poly_polys(lhs: Robj, rhs: List) -> Logicals {
    let n = rhs.len();
    let mut res = Logicals::new(n);
    let xpoly: ExternalPtr<Polygon> = lhs.try_into().unwrap();

    for i in 0..n {
        let ypoly: ExternalPtr<Polygon> = rhs[i].to_owned().try_into().unwrap();
        res.set_elt(i, Rbool::from(xpoly.intersects(&*ypoly)));
    }
    res
}
```


