---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rsgeo

`rsgeo` is an interface to the Rust libraries `geo-types` and `geo`. `geo-types` implements pure rust geometry primitives. The `geo` library adds additional algorithm functionalities on top of `geo-types`. This package lets you harness the speed, safety, and memory efficiency of these libraries. `geo-types` does not support Z or M dimensions. There is no support for CRS at this moment.

```{r}
# remotes::install_github("josiahparry/rsgeo")
library(rsgeo)
```

Create geometries from sf objects

```{r}
# get geometry from sf
data(guerry, package = "sfdep")

polys <- guerry[["geometry"]] |>
  sf::st_cast("POLYGON")

# cast to rust geo-types
rs_polys <- as_rsgeom(polys)

head(rs_polys)
```

Cast geometries to sf

```{r}
as_sf(rs_polys)
```


Calculate the unsigned area of polygons. 

```{r}
bench::mark(
  rust = unsigned_area(rs_polys),
  sf = sf::st_area(polys),
  check = FALSE
)
```
Find centroids 

```{r}
bench::mark(
  centroids(rs_polys),
  sf::st_centroid(polys),
  check = FALSE
)
```
Extract points as matrix

```{r}
rs_polys |> 
  centroids() |> 
  as.matrix() |> 
  head()
```

Plot the polygons and their centroids
```{r}
plot(rs_polys)
plot(centroids(rs_polys), add = TRUE)
```

Calculate a distance matrix 

```{r}
pnts <- centroids(rs_polys)
pnts_sf <- as_sf(pnts)

bench::mark(
  rust = euclidean_distance_matrix(pnts, pnts),
  sf = sf::st_distance(pnts_sf, pnts_sf)
)

```

Simplify a geometry

```{r}
x <- rs_polys[[37]]
x_simple <- simplify_geom(x, 5000)

plot(x)
plot(x_simple, add = TRUE)
```

```{r}
bench::mark(
  rust = simplify_geoms(rs_polys, 500),
  sf = sf::st_simplify(polys, FALSE, 500),
  check = FALSE
)
```

Union geometries with `union_geoms()`

```{r}
plot(union_geoms(rs_polys))
```


Find the closest point to a geometry

```{r}
close_pnt <- closest_point(
  rs_polys[[1]], 
  geom_point(800000, 2090000)
)

plot(rs_polys[[1]])
plot(close_pnt, pch = 15, add = TRUE)
```

Find the haversine destination of a point, bearing, and distance. 

```{r}
bench::mark(
  rust = haversine_destination(geom_point(10, 10), 45, 10000),
  Cpp = geosphere::destPoint(c(10, 10), 45, 10000),
  check = FALSE
)
```


```{r}
origin <- geom_point(10, 10)

destination <- haversine_destination(origin, 45, 10000)

plot(origin)
plot(destination, col = "blue", add = TRUE)
```

Find intermediate point.

```{r}
middle <- haversine_intermediate(origin, destination, 1/2)

plot(origin)
plot(destination, add = TRUE, col = "red")
plot(middle, add = TRUE, col = "blue")
```

Utilize the chaikin smoothing algorithm with 5 iterations.

```{r}
region <- rs_polys[[2]]
plot(chaikin_smoothing(region, 5))
```

Find extreme coordinates with `extreme_coords()`


```{r}
france <- union_geoms(rs_polys)


plot(france)
plot(extreme_coords(france), add = TRUE, pch = 15)
```



#### Notes

Right now plotting is done using `sf` by first casting into R native objects and then assigned the appropriate sf class. That object is then plotted by sf
